---
title: "PA1_template.Rmd"

date: "Friday, August 15, 2014"
output: html_document
---

Loading and preprocessing the data.

```{r, echo = TRUE}
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              "data.zip", quiet = TRUE, mode = "wb")  
unzip("data.zip")

activity = read.csv("activity.csv", header = TRUE) 

```


Eliminate NA from activity.
```{r,echo = TRUE}
newActivity <- activity[!is.na(activity$steps),]

```

Make a histogram of the total number of steps taken each day.

```{r, echo=TRUE}

sumDaily<-aggregate(newActivity$steps, by = list(Date=newActivity$date), FUN =sum)

names(sumDaily)[2] <-"steps"

hist(sumDaily$steps)
```

Calculate and report the mean and median total number of steps taken per day.


```{r, echo=TRUE}

mean(sumDaily$steps)
median(sumDaily$steps)

```

Identify the average daily activity pattern.

```{r,echo = TRUE}

meanInterval<-aggregate(newActivity$steps, by = list(interval=newActivity$interval), FUN =mean)
names(meanInterval)[2] <-"steps"

plot(meanInterval$interval,meanInterval$steps,type="l")


```

Find Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.


```{r,echo = TRUE}
tf<-meanInterval$steps== max(meanInterval$steps)

which(tf)

meanInterval[which(tf),]


```
Imputing missing values:

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

    Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).
    
  
```{r,echo = TRUE}
NAActivity <- activity[is.na(activity$steps),]
nrow(NAActivity)

```

The result is the same as the following:
```{r,echo = TRUE}
nrow(activity)  - nrow(newActivity)
```

Using the mean for that 5-minute interval to fill/impute the mising value.

  
```{r,echo=TRUE}
ImputedActivity<- activity

n.length <- nrow(ImputedActivity)

for (i in 1:n.length)
{
  if (is.na(ImputedActivity[i,1]) )
  {
    s1 <- subset(activity, activity$interval == ImputedActivity[i,3], select = c(steps,interval) )
    ImputedActivity[i,1] <- mean(s1$steps, na.rm = TRUE)
  }

}


```


    With new data set to make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
    
   
```{r, echo=TRUE}
ImputedsumDaily<-aggregate(ImputedActivity$steps, by = list(Date=ImputedActivity$date), FUN =sum)

names(ImputedsumDaily)[2] <-"steps"

hist(ImputedsumDaily$steps)
```

New mean and median for imputed data.

```{r, echo=TRUE}
mean(ImputedsumDaily$steps)
median(ImputedsumDaily$steps)

```

Mean and median differnces.

```{r,echo=TRUE}

meanDifference <- mean(ImputedsumDaily$steps) -mean(sumDaily$steps)
medianDifference <- median(ImputedsumDaily$steps) - median(sumDaily$steps)

meanDifference
medianDifference
```


histogram comparison between Original and Imputed.

Step 1: Adding new type column
Step 2: Combining two dataframes into one.  

```{r,echo = TRUE}

library(ggplot2)

sumDaily$type <-'Original'
ImputedsumDaily$type <- 'Imputed'

df_combined <- rbind(sumDaily, ImputedsumDaily)

ggplot(df_combined, aes(steps, fill = type)) + 
geom_histogram(alpha = 0.5,  position = 'identity')


```

Identify the differences in activity patterns between weekdays and weekends.

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r,echo = TRUE}

weekday <- weekdays(as.Date(ImputedActivity$date, "%Y-%m-%d"))

head(weekday)

for (i in 1:length(weekday)) {
    if ((weekday[i] == "Saturday") | (weekday[i] == "Sunday")) 
        weekday[i] = "weekend" else weekday[i] = "weekday"
}

ImputedActivity$weekday <- as.factor(weekday)

```
    
    
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 


```{r,echo = TRUE}
LevelmeanInterval<-aggregate(ImputedActivity$steps, by = list(interval=ImputedActivity$interval, level=ImputedActivity$weekday), FUN =mean)

names(LevelmeanInterval)[3] <-"steps"

library(lattice)
xyplot(steps ~ interval | weekday, data = ImputedActivity, layout = c(1, 2), type="l", ylab = "number of steps", 
    main = "Average number of steps with two levels: weekday and weekend")

```


